\chapter*{Acknowledgments}
\thispagestyle{noheader}

Foremost, I would like to thank my supervisor \firstTutor{} for the continued support throughout this thesis. The fast and valuable feedback and of course, the possibility to choose a topic of my choice, as well as the taught basics required for writing this thesis really helped me along the way. Also, I have to express my thank-you to \secondTutor, especially for the practice in writing academic works that helped a lot alongside writing this thesis.

Moreover, I thank SinnerSchrader for backing me financially in such a way that I was able to focus my work on this thesis without worries. A personal thank-you is expressed to Ansgar Grapentin for keeping me motivated during this time and of course, for proofreading this thesis.

Further, I would like to acknowledge the valuable feedback from my former co-worker Dr. Ingo Bente.

In addition, I want to thank my fellow students, Gregor, Jasmina, and Jens for always answering my questions about the regularities en route to writing this thesis and for the comprehensive discussions in our study group.

Finally, I would like to express my sincere gratitude to Caro for the countless hours of proofreading, providing useful feedback, and of course the continuous support, motivation, and understanding during this time.

Last but not least, I thank my parents for always supporting me and my decisions, as well as enabling me such a course of studies.

This accomplishment would not have been possible without you all -- thank you.

\newpage