\section{FIDO's Universal Authentication Framework}
\subsection{FIDO Alliance}
\label{subsec:fido_alliance}

The \gls{fido} alliance is an open industry association founded in July 2012 that launched publicly in February 2013. Companies such as PayPal, Lenovo, and Infineon founded the \gls{fido} alliance. Currently, the alliance has more than 260 members, including, e.g., Google, Amazon, Yubico, Samsung, Microsoft, VISA, or MasterCard. The goal of the \gls{fido} alliance is to develop new authentication protocols and standards in order to enhance and simplify the user experience of \gls{mfa} and to reduce the supersaturated usage of passwords. The \gls{fido} alliance developed the specifications \gls{uaf}, \gls{u2f}, \gls{ctap}, and first draft of the \wa.\footcites[See][]{fido-history}[See][583]{eckert-it-sec-9}[See][17]{fido-ct-2}

Another goal of the \gls{fido} alliance is user privacy. As all of their specifications are based on public-key cryptography, this goal is easily achieved. Each key-pair is unique for every registration the user performs and not shared with third-parties. Because of the public-key cryptography, no link between the same user on different websites exists. Further, a \gls{rp}, i.e., web server, is only allowed to access the public-key credentials of a user that are associated the \gls{rp}. Besides, one of the core principles is that biometric data never leaves the local authenticator and that no action is performed without the users' consent. Additionally, no authenticator device is uniquely identifiable. They are only identifiable on a manufacturer or production-batch level.\footcite[See][6--7]{fido-privacy}

\input{content/main_part/multi_factor/fido/uaf}
