The security of the \gls{u2f} protocol extends the security threats of physical security tokens and smart cards that were introduced in the previous section, for example, loss or theft.\footcites[See][12--13]{fido-sec-ref}

In contrast to \gls{otp}, the \gls{u2f} protocol is phishing resistant. This resistance is achieved by binding the token, registration, and authentication to a specific origin. Given the fact that the \gls{u2f} token compares the origin, phishing sites that target, e.g., typos in the \gls{url} do not expose a threat.

However, the \gls{u2f} protocol is not resistant to malware. A malware that controls the \gls{usb} ports of a computer can communicate with the token, too.\footcites[See][10--1]{8429292}[See][9]{u2f-overview}

When using the \gls{tls} ChannelID, also called token binding, the user is protected against \gls{mitm} attacks. The \gls{u2f} protocol specifies the token binding as a possible extension. Token binding enables mutual authentication in \gls{tls} by using cryptographic certificates on both sides. It has to be noted though, that using the \gls{tls} ChannelID is defined as optional in the specification.\footcites[See][6--7]{u2f-overview}

Security researchers were able to find a potential vulnerability in the optional key wrapping of the key material if the counters are not correctly increased and checked. This attack could lead to a successful cloning attack. The security token vendors and the \gls{fido} alliance have adjusted the specifications and recommendations. Further, researchers were able to control the token via the Web\gls{usb} \gls{api} by emulating the \gls{nfc} capabilities of a YubiKey to use it as a smart card instead.\footcites[See][3]{DBLP:journals/corr/abs-1906-06009}

Albeit it is not a vulnerability, the lack of a display on the authenticator can render a threat. An attacker might be able to change the transaction a user signs with the test of user presence. Further, the absence of a display might confuse a user if a registration or authentication was successful. Also, malware can continually request the touch of a button to complete an \gls{mfa} authentication flow with, e.g., phished or key-logged credentials. Due to the lack of a display, the chances are high that a user accidentally confirms a fraudulent authentication from the malware. Additionally, the user might not notice the fraudulent attempt and presses the button again due to the lack of feedback from the token.\footcites[See][434]{10.1007/978-3-662-54970-4_25}[See][15]{das2018johnny}[See][1518--1519]{238325}[See][884]{8418643}

In theory, an attacker could manufacture a backdoor variant of the \gls{u2f} token, which either leaks or sends all the private keys to the attacker. However, this attack has a low probability compared to the other attack vectors.\footcites[See][3]{DBLP:journals/corr/abs-1810-04660}

Moreover, the centralized managed metadata service for device attestation can expose a threat. Given that an attacker takes over this server, they can, e.g., enable counterfeit security tokens to become legitimate attested tokens. On the other hand, legitimate tokens can be flagged as banned or compromised. Besides that, the service is subject to a \gls{ddos} attack that prevents \glspl{rp} from updating their metadata database. Since vendors offer their attestation metadata independently, possible mitigation is to implement an independent metadata service.\footcites[See][]{fido-metadata}[See][17]{Chang2017OnMU}
